---
title: "Identity Fraud Detection DQR"
output: html_notebook
---

Loading the packages and data
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(lubridate)
library(tidyr)
library(scales)
data=read.csv("applications.csv")
summary(data)
```

There are all together 10 variables, with 9 predictors and 1 response (fraud). We will first have some rough understanding of the dataset with summary statistics.
```{r}
# Understand the data type of each variable
str(data)

#Count the number of NAs in each column
t(data%>%
  select(everything())%>%
  summarise_all(funs(sum(is.na(.)))))# There is no missing data in the dataset

#Count the number of unique values in each column
t(data%>%
  select(everything())%>%
  summarise_all(funs(length(unique(.)))))

#Find out the mode for each column
Mode <- function(x) {
  ux <- unique(x[!is.na(x)])
  ux[which.max(tabulate(match(x, ux)))]
}
sapply(data, Mode)
```

After having an overview of the dataset with summary statistics, we start to look at each column specifically with appropriate visualizations.

# SSN
```{r}
## ssn:remove
ssn_top=data %>%
  group_by(ssn) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(21)
#Plot
ggplot(ssn_top,aes(x=reorder(ssn,-count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 SSNs',x='SSN')+
  theme( axis.text.x = element_text(angle = 45, hjust = 1),
         plot.title = element_text(hjust = 0.5),
         text = element_text(size=20))

### remove the most frequent SSN: 737610282
ssn_top=ssn_top[-1,]
ssn_top$ssn=as.factor(ssn_top$ssn)
### plot the top 10 records
ggplot(ssn_top,aes(x=reorder(ssn,-count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 SSNs (Removed Abnormality)',x='SSN')+
  theme( axis.text.x = element_text(angle = 45, hjust = 1),
         plot.title = element_text(hjust = 0.5),
         text = element_text(size=20))
```

# Firstname
```{r}
## firstname
fn_top=data %>%
  group_by(firstname) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(20)
### plot the top 10 records
ggplot(fn_top,aes(x=reorder(firstname,-count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 First Names',x='Firstname')+
  theme( axis.text.x = element_text(angle = 45, hjust = 1),
         plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))
```

# Firstname
```{r}
## lastname
ln_top=data %>%
  group_by(lastname) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(20)
### plot the top 10 records
ggplot(ln_top,aes(x=reorder(lastname,-count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 Last Names',x='Lastname')+
  theme( axis.text.x = element_text(angle = 45, hjust = 1),
         plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))
```

# ZIP5
```{r}
## zip5
zip5_top=data %>%
  group_by(zip5) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(20)
### plot the top 10 records
ggplot(zip5_top,aes(x=reorder(zip5,-count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 Zips',x='Zip')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        text = element_text(size=15))
```

# homephone
```{r}
## homephone: remove the most frequent record: 9105580920
homephone_top=data %>%
  group_by(homephone) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(11)
#Plot
ggplot(homephone_top,aes(x=reorder(homephone,count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 10 Homephone',x='Homephone')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  coord_flip()

#Remove the top 1
homephone_top=homephone_top[-1,]
### plot the top 10 records
ggplot(homephone_top,aes(x=reorder(homephone,count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 10 Homephone (Removed Abnormality)',x='Homephone')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  coord_flip()
```

#date
```{r}
#Convert the date into time format
data$date=mdy(data$date)
data$date=as.character(data$date)
## date
date_top=data %>%
  group_by(date) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(20)
### plot the top 10 records
ggplot(date_top,aes(x=reorder(date,count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 Application Dates',x='Date')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  coord_flip()
## plot the distribution of application for 12 months
ggplot(data,aes(x=as.factor(month(date))))+
  geom_bar(stat = "count", fill='#8da0cb')+
  labs(x='Month',title='Distribution of Applications among 12 Months')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))

# Daily application trend
ggplot(data,aes(x=date))+
  geom_line(stat = "count", color='#8da0cb')+
  labs(title='Application Trend Throughout 2016 (Daily Basis)')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))

# Weekly application trend
ggplot(data,aes(x=as.factor(week(date)),group=1))+
  geom_line(stat = "count", color='#8da0cb')+
  labs(x='Weeks',title='Application Trend Throughout 2016 (Weekly Basis)')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  scale_x_discrete(breaks=seq(1,53,10))

# Weekly application trend
ggplot(data,aes(x=as.factor(month(date)),group=1))+
  geom_line(stat = "count", color='#8da0cb')+
  labs(x='Months',title='Application Trend Throughout 2016 (Monthly Basis)')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  ylim(c(0,9000))
```

#dob
```{r}
## dob
### transform year format
data$dob=as.character(data$dob)

#Split the dob into month, day, and year with only the last two digits
dob_split=data %>%
  separate(dob,into=c('month','day','year_l2'),sep='/') %>%
  select(month,day,year_l2,record)
dob_split$year=as.numeric(dob_split$year_l2)

#add 0 to the years with only one digit. so '7' is tramsformed into '07'
dob_split$year=str_pad(dob_split$year_l2,2,pad='0')
#set all the date to 1900s.
dob_split=mutate(dob_split,year_f2=19)

#Combine all the the year_f2 with year_l2 to get a four digit year column, and then combine all the year, month, and day back to get a transformed dob.
dob_split$year_l2=as.character(dob_split$year_l2)
dob_agg=dob_split %>%
  unite(year,year_f2,year_l2,sep='',remove=FALSE) %>%
  unite(dob_n,month,day,year,sep='/',remove=FALSE)

### merge new dob, year, month with original data dataset 
data=merge(data,dob_agg,by='record')
str(data)
data=data %>%
  select(-year_f2,-year_l2,-day,-month,-year.1,-year)

### change the dob column into time format with lubridate
data$dob_n = as.character(data$dob_n)
data$dob=as.Date(data$dob_n, "%m/%d/%Y")
data=data[,-11]#delete dob_n
data$dob=as.character(data$dob)

#data$date[which(is.na(mdy(data$dob_n)))]
#### Find two observations that failed to parse through lubridate, though the two dobs look good in formate. Probably bug in package. Changed to use as.Date() function.

### plot
dob_top=data %>%
  group_by(dob) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(10)

### plot the top 10 records
library(scales)
ggplot(dob_top,aes(x=reorder(dob,count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 10 Date of Birth',x='Date of Birth')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  scale_y_continuous(trans = log10_trans(),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)))+#scale the x-axis to log form to strech the low bars.
  coord_flip()
```

#text variable: address
```{r}
## address
address_top=data %>%
  group_by(address) %>%
  summarize(count=n()) %>%
  arrange(-count) %>%
  head(20)
### plot the top 10 records
ggplot(address_top,aes(x=reorder(address,count),y=count))+
  geom_bar(stat = 'identity', fill='#8da0cb')+
  labs(title='Distribution of Top 20 Address',x='Address')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))+
  coord_flip()
```

#fraud
```{r}
data$fraud=as.factor(data$fraud)
summary(data$fraud)

ggplot(data,aes(x=fraud,fill=fraud))+
  geom_bar(stat = 'count', fill=c('#8da0cb','#fc8d62'))+
  labs(title='Distribution of Fraud',x='fraud')+
  theme(plot.title = element_text(hjust = 0.5),
         text = element_text(size=15))
```

